---
title: "16S Indicator species"
author: "Lisa-Marie Delpech"
date: "23/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

# Documentation

**Input**: tables from the script **16S_Data_formatting.Rmd** and files / function from the script **16S_Taxonomic_composition.Rmd**
  
**Outputs**:  

*  indicator OTUs for seasonality (other groups *e.g.* water type not relevant due to a gradient, not enough difference between fjord sites and not enough indicators)  
*  filtered indicators based on IV (≥ 0.7) and p-value (≤ 0.001)  
*  tables of highly abundant indicators for seasonal grouping, based on a 0.5% relative abundance threshold, and giving the genus to which belong the OTUs, as well as corresponding number of significant indicator OTUs    
  
**Dependancies**: **indicspecies**, **vegan**, **ggplot2**, **taxa.pooler.1.4.R** (authors: Gobet A. & Ramette A., see: <https://gitlab.leibniz-zmt.de/chh/bioinf/blob/788dc1a3e2a03ba0e60c120e12663e99599ab3e1/taxa.pooler.1.4.r>) 

# Input data and libraries

## Load data tables

```{r}
rarOTU <- readRDS("./R_data/rarOTU.rds")
subsites <- readRDS("./R_data/subsites.rds")
indval.season <- readRDS("./R_data/indval.rds")
taxonomy <- readRDS("./R_data/taxonomy.rds")
subtaxonomy <- readRDS("./R_data/subtaxonomy.rds")
```


## Libraries

```{r}
require(vegan)
require(indicspecies)
require(ggplot2)

source("./Sources/taxa.pooler.1.4.R")
```



# Indicator Value analysis with site group combinations

Compute IndVal using multipatt function from the *indicspecies* package.
By default *multipatt* uses an extension of the original Indicator Value method, because the function looks for indicator species of both individual site groups and combinations of site groups, as explained in De Cáceres et al. [2010]
The option control = how(nperm=9999) allows choosing the number of random permutations required for the permutational test (this number affects the precision of the p-value). Function how from the permute package allows defining more complex permutational designs.
*multipatt* function returns the square root of IndVal.

Data used: rarefied, subsampled


# Indicator species for seasonality

## Indicator value

Run indicator analysis. The **multipatt** function requires a numerical grouping (cluster)  

```{r}
groups.season <- c(as.numeric(subsites$season))
group_names <- subsites$season
```

```{r, eval = FALSE}
indval.season <- multipatt(t(rarOTU),cluster=groups.season,func="IndVal.g",
                    control=how(nperm=9999))

save.RDS(indval.season, "./R_data/indval.rds") # As this step takes time, save the output of multipatt
```

Extract the good information from the analysis  

```{r}
summary(indval.season, indvalcomp=TRUE, alpha=0.001)
IV <- as.data.frame(indval.season$str) # indicator value
sign <- as.data.frame(indval.season$sign)
significance <- data.frame(sign$p.value) # p-values
row.names(significance) <- rownames(sign)
```

Get the significant indicator species  
Filter the indicator species based on indicator value (IV), and significance level (alpha, p-value), for each group independently (ommiting combinations)  

```{r}
IndSpe.season <- list() # 1 Riv, 2 June, 3 Aug, 4 Sed
for (i in 1:length(levels(as.factor(groups.season)))){
  otu <- rownames(IV[which(IV[,i]>=0.7),]) # Filter the IV
  otu <- otu[which(significance[otu,]<=0.001)] # Filter the p-value
  tax <- taxonomy[otu,"genus"] # Get taxonomy accordingly (genus)
  #otu <- otu[- grep("NA_unclassified",tax)] # Omit NA_unclassified and match otu
  #tax <- tax[- grep("NA_unclassified",tax)] # Omit NA_unclassified
  stat <- IV[otu,i] # Get the IV
  alpha <- significance[otu,] # Get the p-value
  IndSpe.season[[i]] <- data.frame(otu,tax,stat,alpha)
}
# output: a list of dataframes with significant indicator OTUs, taxonomy, IV and p value
```

## Output tables of indicators

Get the names of all significant indicators (all groups together)

```{r}
names <- c()
for (i in 1:length(IndSpe.season)){
  names <- c(names,as.character(IndSpe.season[[i]]$otu)) # Collect all indicator OTU names 
}

indicOTU.tax <- subtaxonomy[levels(as.factor(names)),] # Subset taxonomy accordingly
```

Make a taxonomy df for taxa pooler

```{r}
indicOTU.tax2 <- cbind(rownames(indicOTU.tax), indicOTU.tax)
colnames(indicOTU.tax2)[1] <- "OTU"
#rownames(indicOTU.tax2) <- NULL
```

Get the table pooled per genus for all samples (or import from the script **16S_Taxonomic_composition**)

```{r}
Taxa_pooled <- readRDS("./R_data/Taxa_pooled_all.rds")

relGenus_all <- prop.table(t(Taxa_pooled$genus), 2) * 100
```


### Water column in August

```{r}
August.names <- as.character(IndSpe.season[[3]]$otu)
August.indic <- rarOTU[August.names,which(subsites$season == "August")]
dim(August.indic) # 340 x 29 340 indicator otu and 29 samples in August
# Pool OTUs by genus
August.indic.pooled <- taxa.pooler(data.frame(August.indic,indicOTU.tax2[rownames(August.indic),])) # 29, 7, n, n
August.names.genus <- colnames(August.indic.pooled$genus) # 56 indicator genera (high & low)

# Run relative abundance among indicators
Indic_relGenus_Aug <- prop.table(t(August.indic.pooled$genus), 2) * 100

Indic_relGenus_Aug_high <- Indic_relGenus_Aug[apply(Indic_relGenus_Aug, 1, FUN = function(x){mean(x) >= 0.5}),] # Filter based on relative abundance within indicators

rowMeans(Indic_relGenus_Aug_high)
```

Filter based on abundance in August

```{r}
relGenus.indic.August <- relGenus_all[August.names.genus, which(subsites$season == "August")]
August.indic.high <- relGenus.indic.August[apply(relGenus.indic.August, 1, FUN = function(x){mean(x) >= 0.5}),] # The highly abundant indicators are defined as: the mean relative abundance in their group is above 0.5%

Aug.indic.high.names <- rownames(August.indic.high) # 27 taxa
dim(August.indic.high) # number of indicator OTUs 
apply(August.indic.high, 1, function(x){mean(x)}) # proportion of each taxa
sum(apply(August.indic.high, 1, function(x){mean(x)})) # check that this proportion matches mean of sample sum
mean(colSums(August.indic.high)) # = this one
# get the proportion that these taxa explain within the month
sum(apply(data.frame(relGenus_all)[Aug.indic.high.names, which(subsites$season == "August")], 1, function(x){mean(x)})) # 75% (sum of the mean relative abundances for all high indicator taxa)
```

Output a table of high abundant indicators

```{r}
Aug.indic.table <- as.data.frame(apply(data.frame(relGenus_all)[Aug.indic.high.names, which(subsites$season == "August")], 1, function(x){mean(x)})) # proportion of each taxa in August
# or relGenus_s[rownames(Aug.indic.table),"August"] (see 16S_Taxonomic_composition)
colnames(Aug.indic.table) <- "Relative abundance in August"
Aug.indic.table$Indicators <- rownames(Aug.indic.table)
Aug.indic.table
dim(Aug.indic.table) # 28 x 2
```

Get the number of OTUs that makes up each genus  

```{r}
nb.otu.Aug <- as.data.frame(table(t(IndSpe.season[[3]][which(IndSpe.season[[3]]$tax %in% rownames(Aug.indic.table)),1:2])))
row.names(nb.otu.Aug) <- nb.otu.Aug$Var1
nb.otu.Aug

Aug.indic.table$`Number of OTUs` <- nb.otu.Aug[Aug.indic.table$Indicators, "Freq"]
Aug.indic.table
```

Write table to directory

```{r, eval = FALSE}
saveRDS(Aug.indic.table, "./R_data/August_indicators.rds")
```


For further analysis, a higher threshold of abundance is needed (cf. **16S_Envrionmental_drivers**)

```{r}
Aug.indic.table.1 <- Aug.indic.table[which(Aug.indic.table$`Relative abundance in August` >= 1),] # Select indicators that have higher abundance than 1%
Aug.indic.table.1
dim(Aug.indic.table.1) # 21 x 2
```

```{r, eval = FALSE}
saveRDS(Aug.indic.table.1, "./R_data/August_indicators_high.rds")
```

### Water column in June

```{r}
June.names <- as.character(IndSpe.season[[2]]$otu)
June.indic <- rarOTU[June.names,which(subsites$season == "June")] # 204 x 31
June.indic.pooled <- taxa.pooler(data.frame(June.indic,indicOTU.tax2[rownames(June.indic),])) #31, 7, n, n
June.names.genus <- colnames(June.indic.pooled$genus)

# Run relative abundance among indicators
Indic_relGenus_Jun <- prop.table(t(June.indic.pooled$genus), 2) * 100

Indic_relGenus_Jun_high <- Indic_relGenus_Jun[apply(Indic_relGenus_Jun, 1, FUN = function(x){mean(x) >= 0.5}),] # Filter based on relative abundance within indicators

rowMeans(Indic_relGenus_Jun_high)
```

Filter based on abundance

```{r}
relGenus.indic.June <- relGenus_all[June.names.genus, which(subsites$season == "June")]
June.indic.high <- relGenus.indic.June[apply(relGenus.indic.June, 1, FUN = function(x){mean(x) >= 0.5}),]

Jun.indic.high.names <- rownames(June.indic.high) # 16 taxa
dim(June.indic.high) # number of indicator OTUs
apply(June.indic.high, 1, function(x){mean(x)}) # proportion of each taxa
sum(apply(June.indic.high, 1, function(x){mean(x)})) # check that this proportion match mean of sample sum
mean(colSums(June.indic.high)) # = this one
# get the proportion that these taxa explain within the month
sum(apply(data.frame(relGenus_all)[Jun.indic.high.names, which(subsites$season == "June")], 1, function(x){mean(x)})) # 72%
apply(data.frame(relGenus_all)[Jun.indic.high.names, which(subsites$season == "June")], 1, function(x){mean(x)}) # proportion of each taxa in June
```

Output a table of high abundant indicator

```{r}
Jun.indic.table <- as.data.frame(apply(data.frame(relGenus_all)[Jun.indic.high.names, which(subsites$season == "June")], 1, function(x){mean(x)})) # proportion of each taxa in June
colnames(Jun.indic.table) <- "Relative abundance in June"
Jun.indic.table$Indicators <- rownames(Jun.indic.table)
Jun.indic.table
dim(Jun.indic.table) # 16 x 2
```

Get the number of OTUs that makes up each genus  

```{r}
nb.otu.Jun <- as.data.frame(table(t(IndSpe.season[[2]][which(IndSpe.season[[2]]$tax %in% rownames(Jun.indic.table)),1:2])))
row.names(nb.otu.Jun) <- nb.otu.Jun$Var1
nb.otu.Jun

Jun.indic.table$`Number of OTUs` <- nb.otu.Jun[Jun.indic.table$Indicators, "Freq"]
Jun.indic.table
```

Write table to directory

```{r, eval = FALSE}
saveRDS(Jun.indic.table, "./R_data/June_indicators.rds")
```


For further analysis, a higher threshold of abundance is needed (cf. **16S_Envrionmental_drivers**)

```{r}
Jun.indic.table.1 <- Jun.indic.table[which(Jun.indic.table$`Relative abundance in June` >= 1),] # Select indicators that have higher abundance than 1%
Jun.indic.table.1
dim(Jun.indic.table.1) # 15 x 3
```

```{r, eval = FALSE}
saveRDS(Jun.indic.table.1, "./R_data/June_indicators_high.rds")
```

### Rivers

```{r}
River.names <- as.character(IndSpe.season[[1]]$otu)
River.indic <- rarOTU[River.names,which(subsites$season == "River")] #1924 x 11
River.indic.pooled <- taxa.pooler(data.frame(River.indic,indicOTU.tax2[rownames(River.indic),])) # 11, 7, n, n
River.names.genus <- colnames(River.indic.pooled$genus)

# Run relative abundance among indicators
Indic_relGenus_Riv <- prop.table(t(River.indic.pooled$genus), 2) * 100

Indic_relGenus_Riv_high <- Indic_relGenus_Riv[apply(Indic_relGenus_Riv, 1, FUN = function(x){mean(x) >= 0.5}),] # Filter based on relative abundance within indicators

rowMeans(Indic_relGenus_Riv_high)
```

Filter based on abundance

```{r}
relGenus.indic.River <- relGenus_all[River.names.genus,which(subsites$season == "River")]
River.indic.high <- relGenus.indic.River[apply(relGenus.indic.River, 1, FUN = function(x){mean(x) >= 0.5}),] # 35 indicators

Riv.indic.high.names <- rownames(River.indic.high) # 35 taxa
dim(River.indic.high) # number of indicator OTUs
apply(River.indic.high, 1, function(x){mean(x)}) # proportion of each taxa
sum(apply(River.indic.high, 1, function(x){mean(x)})) # check that this proportion matches mean of sample sum
mean(colSums(River.indic.high)) # = this one
# get the proportion that these taxa explain within the month
sum(apply(data.frame(relGenus_all)[Riv.indic.high.names, which(subsites$season == "River")], 1, function(x){mean(x)})) # 70%
apply(data.frame(relGenus_all)[Riv.indic.high.names, which(subsites$season == "River")], 1, function(x){mean(x)}) # proportion of each taxa in the rivers
```

Output a table of high abundant taxa

```{r}
Riv.indic.table <- as.data.frame(apply(data.frame(relGenus_all)[Riv.indic.high.names, which(subsites$season == "River")], 1, function(x){mean(x)})) # proportion of each taxa in the rivers
colnames(Riv.indic.table) <- "Relative abundance in rivers"
Riv.indic.table$Indicators <- rownames(Riv.indic.table)
dim(Riv.indic.table) # 32 indicators 
sum(Riv.indic.table$`Relative abundance in rivers`) # for 70 %
```

Get the number of OTUs that makes up each genus  

```{r}
nb.otu.Riv <- as.data.frame(table(t(IndSpe.season[[1]][which(IndSpe.season[[1]]$tax %in% rownames(Riv.indic.table)),1:2])))
row.names(nb.otu.Riv) <- nb.otu.Riv$Var1
nb.otu.Riv

Riv.indic.table$`Number of OTUs` <- nb.otu.Riv[Riv.indic.table$Indicators, "Freq"]
Riv.indic.table
```

Write table to directory

```{r, eval = FALSE}
saveRDS(Riv.indic.table, "./R_data/River_indicators.rds")
```


For further analysis, a higher threshold of abundance is needed (cf. **16S_Envrionmental_drivers**)

```{r}
Riv.indic.table <- Riv.indic.table[which(Riv.indic.table$`Relative abundance in rivers` >= 1),] # Select indicators that have higher abundance than 1%
Riv.indic.table
dim(Riv.indic.table) # 
```


### Sediments

```{r}
Sed.names <- as.character(IndSpe.season[[4]]$otu)
Sed.indic <- rarOTU[Sed.names,which(subsites$season == "Sediment")]
Sed.indic.pooled <- taxa.pooler(data.frame(Sed.indic,indicOTU.tax2[rownames(Sed.indic),])) # 16 samples
Sed.indic.names <- colnames(Sed.indic.pooled$genus)

# Run relative abundance among indicators
Indic_relGenus_Sed <- prop.table(t(Sed.indic.pooled$genus), 2) * 100

Indic_relGenus_Sed_high <- Indic_relGenus_Sed[apply(Indic_relGenus_Sed, 1, FUN = function(x){mean(x) >= 0.5}),] # Filter based on relative abundance within indicators

rowMeans(Indic_relGenus_Sed)
```

Filter based on a 0.5% relative abundance threshold

```{r}
relGenus.indic.Sed <- relGenus_all[Sed.indic.names, which(subsites$season == "Sediment")]
Sed.indic.high <- relGenus.indic.Sed[apply(relGenus.indic.Sed, 1, FUN = function(x){mean(x) >= 0.5}),] # 40 indicators

Sed.indic.high.names <- rownames(Sed.indic.high) # 40 taxa
dim(Sed.indic.high) # number of indicator OTUs
apply(Sed.indic.high, 1, function(x){mean(x)}) # proportion of each taxa
sum(apply(Sed.indic.high, 1, function(x){mean(x)})) # check that this proportion matches mean of sample sum
mean(colSums(Sed.indic.high)) # = this one
# get the proportion that these taxa explain within the month
sum(apply(data.frame(relGenus_all)[Sed.indic.high.names, which(subsites$season == "Sediment")], 1, function(x){mean(x)})) # 62 %
apply(data.frame(relGenus_all)[Sed.indic.high.names, which(subsites$season == "Sediment")], 1, function(x){mean(x)}) # proportion of each taxa in the sediments
```

Output a table of high abundant indicators

```{r}
Sed.indic.table <- as.data.frame(apply(data.frame(relGenus_all)[Sed.indic.high.names, which(subsites$season == "Sediment")], 1, function(x){mean(x)})) # proportion of each taxa in the sediments
colnames(Sed.indic.table) <- "Relative abundance in sediments"
Sed.indic.table$Indicators <- rownames(Sed.indic.table)
Sed.indic.table <- Sed.indic.table[-37,] # remove unknown family unclassified (6%)
dim(Sed.indic.table) # 39 indicators 
sum(Sed.indic.table$`Relative abundance in sediments`) # for 55 %
```

Get the number of OTUs that makes up each genus  

```{r}
nb.otu.Sed <- as.data.frame(table(t(IndSpe.season[[4]][which(IndSpe.season[[4]]$tax %in% rownames(Sed.indic.table)),1:2])))
row.names(nb.otu.Sed) <- nb.otu.Sed$Var1
nb.otu.Sed

Sed.indic.table$`Number of OTUs` <- nb.otu.Sed[Sed.indic.table$Indicators, "Freq"]
Sed.indic.table
```

Write table to directory

```{r, eval = FALSE}
saveRDS(Sed.indic.table, "./R_data/Sediment_indicators.rds")
```

